---
title: Approximating spatial processes with too many knots degrades predictive accuracy of SPDE models
author: Eric J. Ward$^{1,\dagger}$, Sean C. Anderson$^{2,\dagger}$
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: options.sty
    latex_engine: xelatex
  bookdown::pdf_document2: default
csl: mee.csl
editor_options:
  markdown:
    wrap: sentence
bibliography: how-many-knots.bib
---

$^1$Conservation Biology Division, Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd E, Seattle WA, 98112, USA\
email: [eric.ward\@noaa.gov](mailto:eric.ward@noaa.gov){.email}\
$^2$Pacific Biological Station, Fisheries and Oceans Canada, Nanaimo, BC, V6T 6N7, Canada\
$^\dagger$ Authors contributed equally to this work

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(knitr)
library(ggrepel)
library(viridis)
library(gridExtra)
library(cowplot)
library(jpeg)
```

## Abstract

\setlength\parindent{24pt}

1.  Spatial and spatiotemporal models are increasingly used in ecology.
    Applications have used these methods to track population change over time, assess changing species distributions, and model spatial population processes such as predation, habitat use, and fishing impacts.

2.  Computational advances over the last decade have allowed for the rapid estimation of spatial approximations to high-dimensional spatial surfaces.
    One such advance has been the Stochastic Partial Differential Equation (SPDE) approach that allows for estimating a continuous Gaussian Markov random field that approximates a continuous Gaussian random field.
    One of the most critical decisions for building SPDE-based models is how to construct a finite-element "mesh" that is required to implement the SPDE.
    Little guidance has been established in constructing these meshes, with many previous applications suggesting that higher resolution meshes result in better parameter estimation and/or predictive accuracy.

3.  Using a publicly available dataset of groundfish species on the west coast of the USA, we develop three cases studies.
    These include a spatial model of ocean temperature from a single year, presence-absence models of 27 species over a 21-year period, and delta-models of the same 27 species combining presence-absence with catch rates.
    We use these case studies to understand how mesh resolution affects the predictive accuracy of models, and how parameter inference changes as a function of mesh resolution.

4.  Our spatial model of ocean temperature highlights that the finest scale meshes generally overfit the data and result in decreased out-of-sample predictive performance.
    Consistent with previous work, our results highlight that fixed effect parameters of interest decline in magnitude as the complexity of the spatial field increases.
    Our models of groundfish occurrence and temperature also highlight the relationship between mesh resolution and spatial parameters: as more vertices are added and meshes increase in resolution, estimated spatial range and variance parameters generally decline in magnitude.
    Finally, our applications of delta-models to groundfish demonstrates that for most species, there is little effect of mesh resolution on area-weighted biomass indices that integrate model predictions.

5.  Results from our work highlight that practitioners should not assume the performance of SPDE-based models always improves with higher resolution meshes.
    Instead, each dataset likely has a range of resolutions that results in optimal out-of-sample predictive performance; the resolution of these meshes may be identified using spatial cross-validation approaches, such as the sampling procedures used in our analysis.
    Selecting models with the appropriate spatial complexity is expected to improve the accuracy of parameter estimation and derived quantities (e.g., estimates of population abundance, metrics of distribution) when out-of-sample prediction is a focus of inference.

## Key words

Spatial, spatiotemporal, Stochastic Partial Differential Equation (SPDE), Integrated Nested Laplace Approximation (INLA), Gaussian random field, mesh, predictive accuracy, sdmTMB, cross-validation

## Introduction

Recent advances in computational methods have revolutionized the estimation of spatial and spatiotemporal models in ecology and related fields [@banerjee_bayesian_2012].
These models address key questions, such as tracking species abundance over time and space, identifying environmental predictors of occurrence, modeling disease spread, and quantifying range shifts in distributions [e.g., @ross_accessible_2012; @miller_spatial_2013; @thorson_comparing_2017].
Application of these methods to large georeferenced datasets has become increasingly important as ecological and environmental challenges demand precise and scalable statistical tools.

A range of statistical approaches are available for modeling spatial processes; examples include generalized additive models (GAMs) [@murase_application_2009; @forney_habitat-based_2012; @mgcv], non-parametric approaches [@mutanga_high_2012], and applications of Gaussian random fields [@latimer_hierarchical_2009; @datta_hierarchical_2016].
Estimating these models can be computationally challenging for large georeferenced datasets [@banerjee_hierarchical_2004], necessitating the use of approximations or smoothing.
One solution to approximation is the Stochastic Partial Differentiation Equation (SPDE) approach [@lindgren_explicit_2011].
The SPDE method uses Gaussian Markov random fields (GMRFs) to model the conditional distributions of values as only dependent on neighboring points; thus, rather than estimating a full rank covariance to describe the spatial process as a Gaussian random field, a sparse precision matrix can be estimated [@lindgren_explicit_2011].
Like other approximations, the SPDE approach can be seen as a form of smoothing [@yue2014].
[@miller2020] provide an excellent summary of the recent SPDE literature, and also show the link between these approaches and smoothing splines.

Advances by @rue_approximate_2009 enabled the SPDE approach to be implemented with the Integrated Nested Laplace Approximation (INLA) software as a fast and efficient approximation to estimating high-dimensional Gaussian random fields with the Laplace approximation to the posterior distribution [@lindgren_bayesian_2015].
These models have been extended to include the ability to fit geographic barriers (such as islands or lakes), include anisotropic covariance functions, and non-Gaussian data [@bakka_spatial_2018].
Related software packages to increase the accessibility of these models includes inlabru [@bachl2019] and PointedSDMs [@mostert2023].
As an alternative to INLA, the SPDE approach has also been integrated with Template Model Builder [TMB, @kristensen_tmb_2016] to enable fast marginal maximum likelihood estimation, also via the Laplace approximation.
Software to implement the TMB-SPDE approach includes VAST [@thorson2019], sdmTMB [@anderson2022], and tinyVAST [@tinyVAST] and comparisons between the INLA- and TMB-based approaches can be found in @osgood-zimmerman2022.

Constructing spatiotemporal models with INLA or related software involves many of the same considerations as with simpler non-spatial models.
For example, analysts frequently ask: "Which covariates are most important?" and "Does my model do a satisfactory job of making predictions?" Spatial or spatiotemporal models may include additional considerations, such as specifying the type and complexity of spatial or spatiotemporal variation [@anderson_black_2019].
Because of the flexibility in these approaches, choices about how spatial processes are modeled can also affect inference about covariates.
For example, fitting an overly complex spatial process may mask effects of important covariates [@dambly2023].
One of the more complicated and important decisions---particularly for models implementing the SPDE approach---lies in specifying the dimensionality of the latent spatial fields [@lindgren_bayesian_2015].

A critical yet often overlooked step in SPDE-based modeling is constructing the finite element spatial "mesh"---a triangular discretization that approximates the spatial field (Figure 1).
Functionally, the mesh serves two purposes: (1) it defines the sparse matrices required to implement the SPDE approach and (2) it defines how the spatial values estimated at the vertices (also "nodes" or "knots") are projected to coordinates where data are observed or predictions are made [@lindgren_explicit_2011].
Typically, these vertex values are projected with bilinear interpolation along the mesh triangles [@lindgren_explicit_2011] but piecewise-contestant approximations have also been used [e.g., @thorson2015].
Decisions regarding the mesh, such as its boundary shape, maximum triangle edge length, and minimum triangle angles, significantly impact model accuracy and computational efficiency [@gomez-rubio_bayesian].
While increasing mesh resolution generally improves the approximation of spatial processes, it may also introduce overfitting, computational challenges, and inflated standard errors for predictions.
While many previous applications and guidance around INLA have suggested that increasing the resolution of the mesh improves predictive accuracy [@godana_dynamic_2019; @wilson_evaluating_2021; @williamson_spatiotemporal; @thorson_guidance_2019], a much smaller number of studies have noted that finer meshes do not always translate into better performance [@huang_evaluating_2017] and we are only aware of a small number of analyses examining the importance of mesh construction [@righetto_choice_2020; @roste_2020; @dambly2023].

The objective of this paper is to examine whether higher resolution spatial approximations using the SPDE approach translate to better predictive accuracy, and provide general recommendations for identifying models with good out-of-sample predictive performance.
We develop three case studies based on a publicly available dataset of marine fishes in the Northeast Pacific Ocean.
First, we develop a spatial model of ocean temperature from a single year, investigating the effect of mesh resolution on predictive performance and parameter estimation.
Second, we develop a spatiotemporal model of groundfish occurrence for 27 species, examining the effect of mesh resolution and spatial blocking on predictive performance.
<!-- TODO: come back to check this sentence: EW: I'd like to probably move this off my personal repo and onto an organization? DFO-NOAA-Pacific or unofficial division one, https://github.com/orgs/nwfsc-cb--> Finally, we extend the groundfish models to derive area-weighted population biomass indices, and examine effects of mesh resolution on trend, scale, and standard errors of biomass estimates.
All data for our analysis are publicly available, and code and data to reproduce our analysis are available on Github (<https://github.com/ericward-noaa/how-many-knots>) and Zenodo (on publication).

## Methods

### Data

To evaluate the effect of mesh resolution on the predictive ability of models using the SPDE approach, we used a publicly available scientific survey of marine fishes in the Northeast Pacific Ocean (<https://www.nwfsc.noaa.gov/data>).
The West Coast Groundfish Bottom Trawl Survey (WCGBTS) data represents a fishery-independent survey collected annually over a 21-year period, 2003-2023 (no survey in 2020), covering the California Current ecosystem on the west coast of the USA [@keller_northwest_2017].
The survey is designed to estimate the abundance, sizes, and age composition of commercially and recreationally targeted groundfish species found in near-bottom habitats.
The fundamental methods of the survey, sampling intensity, gear, seasonal and spatial coverage, and random stratified design have remained relatively constant.
In addition to collecting biological samples, the survey also measures environmental variables, such as water temperature at the depth of the net.
Importantly, this survey shares similar characteristics to many other bottom trawl survey datasets used to monitor marine fishes around the world.

### Modeling ocean bottom temperature

As a first case study examining the impacts of mesh resolution on predictive ability and parameter inference, we constructed a spatial model of bottom temperature from the WCGBTS survey.
To keep the initial application simple and avoid complications introduced by spatiotemporal variation, we used a single year of data (2018, n = 701 observations).
Because the survey spans nearly five months (19 May – 15 October), we included day of year as a quadratic predictor, to allow for seasonal changes in mean temperature.
Extending notation used in linear mixed effect models, our model can be expressed as $$
u_{i} = b_{0}+b_{1}x_{i}+ b_{2}x_{i}^{2}+\omega_{i}
$$ where $\omega_{i}$ is the estimated spatial field at the location of the ith observation and modeled as $\mathbf{\omega} \sim MVN(0,\Sigma_\omega)$, $b_0$ is the intercept, $x_{i}$ is the day of year, $b_1$ and $b_2$ are estimated coefficients, and $u_{i}$ is the expected value incorporating fixed and random effects.
Although temperature is strongly correlated with depth, we excluded depth from the model to avoid confounding due to the pronounced spatial gradient along the continental shelf in the California Current system [@keller_northwest_2017].
Finally, we assumed error in the observed temperature values followed a Gaussian distribution.

Mesh resolution was varied by modifying the cutoff distance (representing the distance where two points are treated as a single vertex) because previous research has shown that cutoff distance is one of the more meaningful parameters in determining mesh resolution [@righetto_choice_2020].
There is an inverse relationship between cutoff distance and mesh resolution that varies slightly dataset to dataset; we varied cutoff distances between 8--500km (corresponding to 24--727 vertices for this example).
For each mesh resolution, we performed k-fold cross validation, using 10 folds with points assigned randomly to each.

### Modeling species occurrence

As a second example, we constructed a case study focused on modeling groundfish occurrence from samples taken from the WCGBTS survey, 2003--2023.
The primary objectives of this example are to examine the impacts of mesh resolution on predictive performance in the presence of spatial and spatiotemporal variation, and to explore cross validation designs that involve spatial blocking.
We selected 27 species of commercial or conservation interest, representing a range of occurrence rates (2--60% of hauls) and spatial ranges, driven by differences in life history and movement (Table 1).

We applied the same form of the spatiotemporal model to all species so that for observation $i$, $$
g(u_{i}) = \mathbf{X_{i}}\mathbf{b}+ \omega_{i} + \epsilon_{i,t}
$$ where for convenience the fixed effect design matrix and coefficients are collapsed into $\mathbf{X}$ and $\mathbf{b}$ respectively, $\omega_{i}$ represents spatial variation $\mathbf{\omega} \sim \operatorname{MVN}(0,\Sigma_\omega)$ and $\epsilon_{i,t}$ represents spatiotemporal variation that is assumed independent by year, $\mathbf{\epsilon_{t}} \sim \operatorname{MVN}(0,\Sigma_\epsilon)$.
We modeled presence--absence with a binomial distribution, and $g()$ represents a logit link function.
The only covariate included in the model was year (as a factor), allowing the average probability of occurrence to vary by year.

We expanded on our initial case study and developed custom meshes.
First, we used a non-convex hull to define a boundary, and incorporated the boundary into each mesh, along with a specified cutoff distance (12--175km).
We used 5% offsets for the mesh and boundary (shrinking these toward the data) and set the maximum inner and outer triangle edges at 100 and 500km, respectively.
Meshes varied slightly by species, but the range of cutoff values resulted in meshes ranging in size from 51--893 vertices.
High-resolution meshes were included as an upper extreme, though many applications of the SPDE approach to datasets with similar dimensions have used 500--600 knots [e.g., @thorson_guidance_2019; @tolimieri_spatio-temporal_2020].

For each combination of species and mesh, we performed k-fold cross validation, using 10 folds and random fold assignments, as in our temperature case study.
To better understand the impact of spatial blocking in assigning folds, we also performed cross-validation using spatial blocks.
Given the rectangular domain of the survey area (Figure 1), we divided the domain into latitudinal strips with a fixed strip width, and assigned each resulting strip to a fold in ascending order (1, 2, 3, ..., 8, 9, 10, 1, 2, ...).
We explored a range of strip heights for each species (from 10 to 150km, in steps of 20km) and repeated the model fitting for each.

### Modeling biomass indices

As a third example, we examined the effect of mesh resolution on biomass indices generated from the WCGBTS survey data.
A common use of bottom trawl survey data is to estimate annual indices of biomass, to be used in downstream integrated population models (stock assessments in fisheries).
We used data from the same 27 species from our occurrence modeling example (Table 1) to generate standardized indices of biomass [@thorson2015; @thorson2013].
For each species, we constructed three meshes with cutoff distances of 15, 50, and 100km.
Unlike our previous examples, we did not apply cross validation, but fit a single model to the entire dataset.
For simplicity, we assumed the mesh for each species covered the entire spatial domain of the survey, and we did not filter data a priori (some previous population models have applied depth filters for instance).
Catch per unit effort (CPUE, in kg per km2) was used as a response variable, and because observations are both zero-inflated and skewed, we modeled CPUE as a two-part hurdle- or delta-model, with a binomial model used for presence–absence and a Gamma distribution used for positive catches [@pennington1983].
Like our occurrence models, we only included year effects (as a factor variable) – though in this application, two sets of fixed effects coefficients are estimated, one for each sub-model.
We also included spatial fields and spatiotemporal fields (assumed independent by year), with a shared range assumed between the spatial and spatiotemporal fields [@anderson2022].
Standardized indices of biomass were generated by making model predictions on the survey design grid for each year (12000 2nm by 1.5nm cells of equal area; [@keller_northwest_2017]) and then summing across grid cells within a year to get annual indices of total biomass.

### Estimation

We fit all models in R [@r2023] using the package sdmTMB [@anderson2022], which combines the SPDE approach for spatial models with the speed of maximum marginal log likelihood estimation using Template Model Builder (TMB, [@kristensen_tmb_2016]).
TMB leverages automatic differentiation to quickly converge on parameter estimates, and it applies the Laplace approximation to integrate out random effects, including spatial or spatiotemporal components.
We used the `sdmTMB_cv()` function to implement k-fold cross validation for all examples.
Model convergence was assessed using the `sanity()` function in sdmTMB; we only retained models that successfully converged (positive-definite Hessian matrix, maximum absolute log likelihood gradient \< 0.001).

### Quantifying performance

To quantify the predictive performance of models with different mesh configurations, we calculated two measures of performance for each of our first two examples and applied them separately to training and test data in each fold.
For a first metric, we calculated Root Mean Squared Error (RMSE) as a measure of accuracy for the temperature modeling example, and the Area Under the Curve (AUC) statistic from the Receiver Operating Characteristic (ROC) occurrence for our presence absence models of groundfish.
Second, we calculated the total log-predictive density (log-score), which evaluates both accuracy and precision.

## Results

### Temperature case study

Our spatial models of ocean temperature demonstrated that the relationship between cutoff distance (or vertices in the mesh) and predictive density was generally in opposite directions for training data used to fit the model versus out-of-sample testing data used to evaluate predictions (Figure 2).
Predictive density increased for the training data (indicating an improvement in fit) as the mesh resolution increased (Figure 2), and increased for the test data up to 400 vertices.
For higher resolution meshes, our results demonstrate a decline in predictive performance for out of sample data.

Examining the relationship between estimated model parameters and cutoff distance (or vertices in the mesh), we found that fixed effect parameters associated with the day of year were largest in magnitude for coarse meshes, and became closer to 0 as the number of vertices increased (Figure 3).
Our results also highlight that the highest resolution meshes are associated with the smallest values of spatial range, spatial variability, and observation error variance (Figure 3).
As meshes become more coarse, observation error variance increases, and the spatial parameters appear to increase initially before stabilizing (Figure 3).

### Groundfish occurrence

For each of the 27 species in our analysis, we fit a total of 64 models with spatial cross validation.
To facilitate comparisons across species, we further filtered the results to focus on the 20 species with models that passed the most convergence criteria (\> 22 of 64 models).
Results from these occurrence models appeared to be largely consistent with our temperature models, in that estimated spatial ranges generally become smaller with finer meshes, and there is consistency across bin widths used for spatial blocking (Figure 4).
Like our temperature example, spatial variance did appear to decrease as meshes became finer for a number of species (sablefish, lingcod, Figure S2).
An interesting effect of adding vertices and making more high resolution meshes is that estimates of spatiotemporal variation became larger for a number of species (Figure S3).
A likely explanation for this pattern is that finer meshes allow the spatial field to capture smaller-scale variability -- as the spatial variance declines, a larger portion of total variance becomes redistributed to spatiotemporal variability (Figure S4).
For a handful of species, we found problematic cases where the estimated spatial range was smaller than the cutoff distances used (aurora, rosethorn, greenstriped rockfishes; Figure 4).

Results from our occurrence models are also useful in contrasting cross validation with random fold assignments to cross validation with spatial blocking.
For all species, we found that random fold assignments were associated with the highest predictive log-likelihood for held out test data and these log-likelihoods were highest for the highest resolution meshes (Figure 5).
We found that as strip widths of spatial blocks increased, the predictive log-likelihood generally declined.
The mechanism responsible for this pattern is that as the bin size increases, our spatial models are trying to predict further and further beyond the range of data (in some cases, exceeding the spatial range) and predictive ability becomes greatly diminished.
As cutoff distances increased relative to strip widths, we found generally less differences in predictive performance across models, because the range of strip widths examined was generally within the cutoff distance (Figure 5).

### Groundfish biomass indices

Our spatiotemporal models of groundfish biomass included year effects, but no other covariates.
For many species in our dataset, we found similar trends and biomass indices across the three cutoff distances examined (15, 50, 100km; Figure S1).
In some cases, the scale of trend estimates was shifted slightly higher or lower (aurora rockfish, longspine thornyhead), but the average standard errors of biomass estimates are qualitatively similar and do not depend on cutoff distances.
The exception to this general pattern was that for several rockfish species, both the biomass scale and standard errors did appear to be impacted by the choice of cutoff distance (greenspotted, greenstriped, rosethorn, sharpchin, stripetail; Figure 5).

## Discussion

Spatial and spatiotemporal models are becoming increasingly used in ecology, fisheries, and related sub-fields, in particular to address questions related to changing species distributions in time and space.
High dimensional datasets with hundreds of observations introduce computational challenges, but complicated spatial structure can be efficiently approximated using the SPDE approach [@miller2020].
Combined with the INLA software for estimation [@rue_approximate_2009], these tools have become widely accessible.
Though several examples and best practices papers have been written to demonstrate how to construct models or specify hyperparameters [@lindgren_bayesian_2015; @gomez-rubio_bayesian], advice on specifying the SPDE mesh to approximate spatial processes has been limited.

In some situations, fitting SPDE models with higher resolution meshes is advantageous.
Finer meshes provide a better approximation of the latent spatial process by capturing smaller-scale variation and improving model fit to training data.
Consistent with previous work [@roste_2020], our results show that increasing mesh resolution up to a point stabilizes parameter estimates (Figure 3).
However, as resolution increases, the spatial fields gain flexibility and may begin to absorb variability that could otherwise be explained by fixed effects.
This redistribution of variability can obscure the relationships between covariates and the response, particularly when the scale of the covariates aligns poorly with the scale of the mesh.
For studies focused on identifying precise relationships between environmental covariates and species' occupancy or density, we recommend exploring a range of mesh resolutions, as finer meshes may not always be most appropriate.
For applications related to index standardization, where the goal is often to derive annual indices of biomass, fitting the highest resolution meshes that are computationally feasible is reasonable for most species.
A caveat from our results, is that for a small number of species examined, the choice of mesh resolution did impact point estimates and variability of those estimates.

Aside from predictive performance, there are several potential downsides to fitting models with high resolution meshes.
First, overly complex spatial fields may dominate potential fixed effects of interest, making coefficients for the latter smaller in magnitude.
Our spatial model of temperature highlights that seasonal effects become smaller for higher resolution meshes, consistent with findings from similar modeling approaches [@dambly2023].
Second, our results highlight a trade-off between mesh complexity and estimated spatial parameters.
As the the spatial field becomes more complex, it can better fit localized spatial variation (otherwise explained by fixed effects, observation error, or other sources of variability), resulting in smaller estimates of spatial range and variance.
As spatial variances becomes smaller, variability may be redistributed into the spatiotemporal variance component.

While there are several alternative ways to measure model performance, we recommend using out of sample methods including k-fold cross-validation to quantify predictive accuracy across a range of meshes.
Tools such as AIC [@akaike1973information] have also been used to discriminate among these types of models; however, AIC is known to be problematic when used as a metric to compare mixed effects models [@harrison_brief_2018] or as a tool to estimate spatial complexity [@velasco_2019].
<!-- TODO: While both marginal and conditional variants of AIC exist (the former using the marginal likelihood integrating out the random effects, the latter using the likelihood conditioned on random effects), these statistics may not appropriately penalize complexity, and may lead to different rankings of models. Other metrics, including the Widely Applicable Information Criterion (WAIC [@watanabe2013]) have been proposed based on Bayesian methods. ... --> While WAIC is available in INLA, higher resolution meshes generally correspond to better WAIC values [@roste_2020].

Like with mesh resolution, we recommend taking an iterative approach to exploring the size of spatial blocks used when applying spatial cross-validation.
Spatial blocks that are too large can result in test sets that are effectively beyond the spatial range of the training data, leading to overly pessimistic assessments of model performance.
Conversely, if blocks are too small, the training and test sets may become highly correlated, inflating performance metrics and leading to biased parameter estimates [@valavi2018].

Comparing the performance of models with alternative spatial complexity introduces additional computational burdens, both in performing cross-validation, and comparing models with different numbers of knots.
This challenge is not unique to models using the INLA-SPDE approach; selecting both the location and number of knots is a general problem for knot-based spatial models.
Other examples of spatial statistics models that have encountered this issue include machine learning methods [reviewed in @banerjee2013], Gaussian Process models [@garton2020] and P-splines [@ruppert2002].
Identifying a range of knot values that yield good predictive performance will be specific to individual datasets (e.g., magnitude of observation error, spatial autocorrelation, number of observations, type of response).
The optimal mesh resolution for one species may be informative for other species from the same dataset---potentially reducing the computational overhead.
Though some work has suggested vertices be chosen relative to the estimated spatial range of a dataset [@roste_2020], this advice creates a somewhat circular framework in that the range is usually known after a reasonable spatial model is constructed.

If these tools are being used to compare models with different covariates, or to compare spatial parameters (ranges, variances) across models, it is critical to ensure that model performance is not confounded by arbitrary choices in mesh resolution.
We recommend fitting models across a range of mesh resolutions to identify those that balance goodness of fit to the training data with predictive performance (out-of-sample evaluation, such as cross-validation).
When using cross-validation with spatial blocking, we suggest choosing block (or strip) dimensions smaller than or comparable to the estimated spatial range.
Blocks that are too large can bias predictions by placing test data outside the spatial range of the training set.
Conversely, blocks that are too small may exaggerate model performance by creating strong correlations between training and test data.
Where spatial parameters or covariate effects are being compared across models (to infer biological or environmental drivers for example), consistent mesh specifications – or at minimum, sensitivity analyses to mesh complexity – should be used to avoid drawing spurious conclusions based on modeling artifacts.
While such sensitivities add computational overhead, they will ensure the robustness and interpretability of statistical inference from these spatial models.

## Acknowledgements

We would like to express our gratitude to Dave Miller, Andy Seaton, and Ole Shelton for helpful comments that improved our analysis and presentation of results.

\break

## Figures

```{r fig1, echo=FALSE, fig.cap = "Example of two meshes constructed for the California Current Ecosystem, using the 2018 WCGBTS survey data. Cutoff values of 80 and 20 km are used, translating into 110 and 394 knots (vertices) respectively. The blue polygon illustrates the boundary between the inner and outer mesh where the outer mesh has a larger maximum triangle edge length.", out.width="70%", fig.align="center"}
knitr::include_graphics("figures/Figure_01.pdf")
```

\break

```{r fig2, echo=FALSE, fig.cap = "(a) Sampled ocean temperature from the 2018 WCGBTS survey. Sample locations further offshore are typically colder and at greater depth. (b--c) Average point-wise log-likelihood as a function of mesh features (b) cutoff distance and (c) number of vertices, where the log-likelihood is calculated for training and test datasets and normalized by the number of samples in each using 10-fold cross validation with random fold assignments.", out.width="5in", fig.align="center"}
knitr::include_graphics("figures/Figure_02.pdf")
```

\break

```{r fig3, echo=FALSE, out.width="5in", fig.align="center", fig.cap = "Estimated fixed effect parameters (top row) and spatial parameters (bottom row) as a function of mesh vertices (more vertices translates to a smaller cutoff distance in creating the mesh). The panels 'day' and 'day2' represent the linear and quadratic effects of day of year. Solid lines represent mean estimates, and shaded ribbons represent 95% confidence intervals."}
knitr::include_graphics("figures/Figure_02b.pdf")
```

\break

```{r fig4, echo=FALSE, out.width="6in", fig.align="center", fig.cap = "Estimated spatial range as a function of mesh vertices, for 20 groundfish species included in our occurrence modeling (the mesh cutoff distance corresponding to each vertex is shown as a solid black line). Colored points represent different strip widths used for latitudinal spatial block cross validation, and dark grey points represent results from cross validation with random fold assignment."}
knitr::include_graphics("figures/groundfish_range_v_n.pdf")
```

\break

```{r fig5, echo=FALSE, out.width="6in", fig.align="center", fig.cap = "Estimated predictive log-likelihood as a function of mesh vertices (more vertices translates to a smaller cutoff distance in creating the mesh), for 20 groundfish species included in our occurrence modeling. Colored points represent different strip widths used for latitudinal spatial block cross validation, and dark grey points represent results from cross validation with random fold assignment."}
knitr::include_graphics("figures/groundfish_loglik_v_vertices.pdf")
```

\break

```{r fig6, echo=FALSE, out.width="5in", fig.align="center", fig.cap = "Mean standard errors of biomass estimates (log space)  estimated for each species across three mesh resolutions (larger cutoff distances result in coarser meshes)."}
knitr::include_graphics("figures/mean_index_se.pdf")
```

```{=tex}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{section}{0}
\setcounter{subsection}{0}
\setcounter{subsubsection}{0}
\resetlinenumber
\setcounter{page}{1}
\setcounter{equation}{0}
```
\break

```{r fig_s1, echo=FALSE, out.width="6.5in", fig.align="center", fig.cap = "Time series of biomass estimates estimated for each species, across three mesh resolutions (larger cutoff distances result in coarser meshes)."}
knitr::include_graphics("figures/SI_figure_indices.pdf")
```

\break

```{r fig_s2, echo=FALSE, out.width="6.5in", fig.align="center", fig.cap = "Estimated spatial standard deviation as a function of mesh vertices (more vertices translates to a smaller cutoff distance in creating the mesh), for 20 groundfish species included in our occurrence modeling. Colored points represent different strip widths used for latitudinal spatial block cross validation, and dark grey points represent results from cross validation with random fold assignment."}
knitr::include_graphics("figures/groundfish_sigmaO_v_vertices.pdf")
```

\break

```{r fig_s3, echo=FALSE, out.width="6.5in", fig.align="center", fig.cap = "Estimated spatiotemporal standard deviation as a function of mesh vertices (more vertices translates to a smaller cutoff distance in creating the mesh), for 20 groundfish species included in our occurrence modeling. Colored points represent different strip widths used for latitudinal spatial block cross validation, and dark grey points represent results from cross validation with random fold assignment."}
knitr::include_graphics("figures/groundfish_sigmaE_v_vertices.pdf")
```

\break

```{r fig_s4, echo=FALSE, out.width="6.5in", fig.align="center", fig.cap = "Ratio of estimated spatial standard deviation to estimated spatiotemporal standard deviation as a function of mesh vertices (more vertices translates to a smaller cutoff distance in creating the mesh), for 20 groundfish species included in our occurrence modeling. Colored points represent different strip widths used for latitudinal spatial block cross validation, and dark grey points represent results from cross validation with random fold assignment."}
knitr::include_graphics("figures/groundfish_sigmaRatio_v_vertices.pdf")
```

\break

```{r table1, echo=FALSE}
species <- read.csv(here::here("species_names.csv"))
species$sci_name <- paste0("*", species$sci_name, "*")
knitr::kable(species, caption = "Table of groundfish species included in our encounter probability modeling.", col.names = c("Common name", "Scientific name", "Proportion of positive tows"))
```

\break

# References
