---
output:
  pdf_document:
    keep_tex: yes
---

  
```{r global_options, include=FALSE}
# Knitr options
knitr::opts_chunk$set(fig.width=3, fig.height=2, 
                      fig.path='figures/',
                      size="footnotesize",
                      echo=TRUE, warning=FALSE, message=FALSE)
```

# Load inlabru

```{r}
library(INLA)
library(inlabru)
```
# Set default INLA values for tutorial session and make shortcuts to a nicer colour scale:

```{r}
init.tutorial()
colsc <- function(...) {
  scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(11,"RdYlBu")),
                       limits = range(..., na.rm=TRUE))
}
colsc_c <- function(...) {
  ggplot2::scale_color_gradientn(colours = rev(RColorBrewer::brewer.pal(11,"RdYlBu")),
                                 limits = range(..., na.rm=TRUE))
}
```

# Create survey region and mesh, and plot mesh:
```{r mesh, fig.width=8, fig.height=8}
bnd <- spoly(data.frame(lon = c(0, 10, 10, 0), lat = c(0, 0, 10, 10)),
             crs=inla.CRS("longlat"))
mesh_fine <- inla.mesh.2d(boundary = bnd, max.edge = 0.2)
ggplot() + gg(mesh_fine)
```

# Create simulated field.
Note: the priors here will not be used in estimation
```{r}
matern_fine <-
  inla.spde2.pcmatern(mesh_fine, 
                      prior.sigma = c(1, 0.01), 
                      prior.range = c(1, 0.01))
true_range <- 4
true_sigma <- 1
true_Q <- inla.spde.precision(matern_fine, theta=log(c(true_range, true_sigma)))
```
What is the pointwise standard deviation of the field? Along straight boundaries,
the variance is twice the target variance. At corners the variance is 4 times as large.

# Generate a sample from the model (use argument "seed" for reproducibility):

```{r}
true_field <- inla.qsample(1, true_Q, seed = 12345L)[,1]

truth <- expand.grid(x=seq(0, 10, length=100),
                     y=seq(0, 10, length=100))
truth$field <- inla.mesh.project(mesh_fine,
                                 loc = as.matrix(truth),
                                 field = true_field)
coordinates(truth) <- c("x", "y")
truth <- as(truth, "SpatialPixelsDataFrame")
```

# Generate sample points and sample from field

```{r}
n <- 200
set.seed(123L)
geosamp <- data.frame(x = runif(n, 0, 10), y = runif(n, 0, 10))
geosamp$observed <- inla.mesh.project(mesh_fine,
                                     loc = as.matrix(geosamp),
                                     field = true_field) + rnorm(n, sd = 0.4)
coordinates(geosamp) <- c("x", "y")
```

# Plot the observed values

```{r sample, fig.width=5, fig.height=5}
ggplot() +
  gg(geosamp, mapping = aes_string("x", "y", colour="observed")) +
  coord_equal() +
  ggtitle("Observed values") +
  colsc_c(geosamp$observed)
```

# Estimating the field

```{r}
mesh <- inla.mesh.2d(boundary = bnd, max.edge = 0.5)
matern <-
  inla.spde2.pcmatern(mesh, 
                      prior.sigma = c(10, 0.01), 
                      prior.range = c(1, 0.01))
cmp <- observed ~ field(map = coordinates,
                        model = matern) +
  Intercept

fit <- bru(cmp, geosamp, family = "gaussian")
```

# Predict the field on a lattice, and use the function generate() to sample from the posterior 

```{r}
pix <- pixels(mesh, nx = 200, ny = 200)
pred <- predict(fit, pix,
                ~ field + Intercept)
samp <- generate(fit, pix,
                 ~ field + Intercept,
                 n.samples = 1)
pred$sample <- samp[[1]]
```

# Plot the true field with sample point locations superimposed
```{r plotsurface, fig.width=12, fig.height=4}
pl_truth_samp <- ggplot() +
  gg(truth, mapping=aes_string(x="x", y="y")) +
  theme(legend.position = "none") +
  gg(geosamp, shape="+", cex=2.5) +
  coord_equal() +
  ggtitle("(a) True field") +
  coord_fixed()
# Plot fitted mean
pl_posterior_mean <- ggplot() + 
  gg(pred, mapping=aes_string(x="x", y="y")) + 
  theme(legend.position = "none") +
  gg(bnd) +
  ggtitle("(b) Posterior mean") + 
  coord_fixed()
# Plot posterior sample
pl_posterior_sample <- ggplot() + 
  gg(pred, mapping=aes_string(x="x", y="y", fill="sample")) + 
  theme(legend.position = "none") +
  gg(bnd) +
  ggtitle("(c) Posterior sample") + 
  coord_fixed()

# Common colour scale for the truth and estimate:
csc <- colsc(truth$field, pred$mean, pred$sample)
multiplot(pl_truth_samp + csc,
          pl_posterior_mean + csc,
          pl_posterior_sample + csc,
          cols = 3)
```
