---
output:
  pdf_document:
    keep_tex: yes
---


```{r global_options, include=FALSE}
# Knitr options
knitr::opts_chunk$set(fig.width=3, fig.height=2, 
                      fig.path='figures/',
                      size="footnotesize",
                      echo=TRUE, warning=FALSE, message=FALSE)
```

# Load inlabru and gorilla data

```{r}
library(INLA)
library(inlabru)
library(RColorBrewer)
data(gorillas)

gnests <- gorillas$nests
mesh <- gorillas$mesh
boundary <- gorillas$boundary
gcounts <- gorillas$plotsample$counts
plots <- gorillas$plotsample$plots
plotnests <- gorillas$plotsample$nests
```

Plot the data

```{r}
fig.plots <- ggplot() + 
  gg(boundary) +
  gg(plots) + 
  gg(gnests, size=0.07,color = "blue") +
  gg(plotnests,size=0.07,color = "red") +
  geom_label(aes(label=gcounts$count, x=gcounts$x+0.35, y=gcounts$y-0.3),
             fill = "white", size = 1.5, label.padding = unit(0.1, "lines")) + 
  coord_fixed() + xlab("x") + ylab("y")
fig.plots
```

# Inference using Poisson count likelihood

Define random variables

```{r}
constr <- FALSE
mdl <- count ~ spat(map = coordinates, 
                    model = inla.spde2.matern(mesh, constr = constr)) +
  Intercept
```

The `constr = TRUE` parameter imposes an integrate-to-zero constraint on the random field to make the model more identifiable.

Fit the model

```{r}
fit <- bru(mdl, family = "poisson", data = gcounts,
           options = list(E = gcounts$exposure))
```

# Results

Predict the intensity

```{r}
pxl <- pixels(mesh, mask = boundary, nx = 500, ny = 400)
dens <- predict(fit, pxl, ~ exp(spat+Intercept))
```

Plot it

```{r}
fig.brudens <- ggplot() + 
  gg(dens) + 
  gg(boundary, alpha = 0) + 
  gg(plots) + 
  scale_fill_gradientn(colours = brewer.pal(9,"YlOrRd"),
                       limits = c(0,max(dens$mean))) +
  gg(gcounts, mapping = aes(size = count)) +
  scale_size_area(max_size = 2)
fig.brudens
```


# Inference using LGCP likelihood (plot data)

Define the model

```{r}
mdl <- coordinates ~ spat(map = coordinates, 
                          model = inla.spde2.matern(mesh, constr = constr)) +
  Intercept
```

Fit it

```{r}
init.tutorial() # Sets Empirical Bayes option, "eb", for speed.
fit <- lgcp(mdl, plotnests, samplers = plots)
```

# Results


```{r}
lgcp.dens <- predict(fit, pxl, ~ exp(spat+Intercept))
fig.lgcpdens <- ggplot() + 
  gg(lgcp.dens) + 
  gg(boundary, alpha = 0) + 
  gg(plots) + 
  scale_fill_gradientn(colours = brewer.pal(9,"YlOrRd"),
                       limits = c(0,420))
fig.lgcpdens
```

# Inference using LGCP likelihood (all data)

Fit the model (useing the same random variables as above)
```{r}
fit.all <- lgcp(mdl, gnests, samplers = boundary)
```

# Results

Predict and plot the intensity

```{r}
lgcp.dens.all <- predict(fit.all, pxl, ~ exp(spat+Intercept))
fig.lgcpdens.all <- ggplot() + 
  gg(lgcp.dens.all) + 
  gg(boundary, alpha = 0) + 
  gg(plots) + 
  scale_fill_gradientn(colours = brewer.pal(9,"YlOrRd"),
                       limits = c(0,420))
fig.lgcpdens.all
```

# Generate plots as shown in the paper

```{r}
noyticks <- theme(
  axis.text.y = element_blank(),
  axis.ticks = element_blank())

noxticks <- theme(
  axis.text.x = element_blank(),
  axis.ticks = element_blank())

colsc.bru <- scale_fill_gradientn(colours = brewer.pal(9, "Oranges"),
                                  limits = c(0, max(dens$mean)))
colsc <- scale_fill_gradientn(colours = brewer.pal(9, "Oranges"),
                              limits = c(0, 420))
```


```{r gnest_counts}
fig.plots + noyticks + noxticks + ylab("") + xlab("(a)")
```


```{r gnest_bru}
fig.brudens + noyticks + noxticks + ylab("") + xlab("(b)") + colsc.bru +
  guides(fill = FALSE, size = FALSE) + coord_equal()
```


```{r gnest_lgcp}
fig.lgcpdens + noyticks + noxticks + ylab("") + xlab("(c)") + colsc +
  guides(fill=FALSE) + coord_equal()
```

```{r gnest_lgcp_all}
fig.lgcpdens.all + noyticks + noxticks + ylab("") + xlab("(d)") + colsc +
  guides(fill=FALSE) + coord_equal()
```

