---
title: SPDE meshes and degraded predictions at high resolutions (with TMB)
author: Sean Anderson, Eric Ward
---

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(sdmTMB)
library(dplyr)
library(future)
# plan(sequential)
plan(multisession)
```

Simulate some data:

```{r sim-data, cache=TRUE}
x <- seq(-1, 1, length.out = 150)
y <- seq(-1, 1, length.out = 150)
loc <- expand.grid(x = x, y = y)
time_steps <- 1
N <- nrow(loc)
X <- model.matrix(~ x1, data.frame(x1 = rnorm(N * time_steps)))
mesh <- make_mesh(loc, xy_cols = c("x", "y"), cutoff = 0.1)
plot(mesh$mesh, asp = 1, main = "")
s <- sdmTMB_sim(
  x = loc$x, y = loc$y, mesh = mesh, X = X,
  betas = c(0.2, 0.3), time_steps = time_steps,
  phi = 0.2, thetaf = 1.3, range = 0.2, sigma_O = 0.4,
  seed = 1234, family = gaussian()
)
head(s)
```

Look at simulated data.

True response:

```{r}
ggplot(s, aes(x, y, fill = mu))+ 
  geom_raster() +
  scale_fill_viridis_c()
```

True spatial random field:

```{r}
ggplot(s, aes(x, y, fill = omega_s))+ 
  geom_raster() +
  scale_fill_gradient2()
```

Sample 500 points to observe:

```{r}
set.seed(1029)
d <- s[sample(seq_len(nrow(loc)), 200), ]
ggplot(d, aes(x, y, colour = observed, size = abs(observed))) + 
  geom_point() +
  scale_color_gradient2() +
  scale_size_area()

# select folds
k_folds = 8
d$fold_ids = sample(1:k_folds, size=nrow(d), replace=T)
```

For different mesh sizes, look at the number of predictions that are being made in triangles with no training data. 

```{r tmb-sim-fits, cache=TRUE, message=FALSE, warning=FALSE}
source("code/archive/15_triangles.R")

cutoffs <- c(0.5, 0.4, 0.3, 0.2, seq(0.19,to=0.05,by=-0.01))
p_new_tri <- 0
for(i in 1:length(cutoffs)) {
  mesh <- inla.mesh.2d(
    loc = d[,c("x","y")],
    #offset = c(-0.05, -0.05),
    cutoff = cutoffs[i],
    max.edge = c(1,10)
  )
  d$new_triangle <- 0 # indicator to hold whether data point is in triangle that isn't in training set
  
  for(f in 1:k_folds) {
    # for each mesh, calculate triangles in the training set
    train_tri <- overlay(mesh, d[which(d$fold_ids!=f),c("x","y")])
    test_tri <- overlay(mesh, d[which(d$fold_ids==f),c("x","y")])
    # get triangle ids with data
    train_names <- names(table(train_tri$id))
    test_names <- names(table(test_tri$id))
    # figure out which polygons are being projected out of the training set
    new_tri <- which(test_names %in% train_names == FALSE)
    # find the ids of individual data points in those test set triangles
    mismatch = which(test_tri$id %in% test_names[new_tri])
    # add indicator for these mismatches
    d$new_triangle[which(d$fold_ids==f)[mismatch]] = 1
  }
  p_new_tri[i] = mean(d$new_triangle)
}


```

Fit an 8-fold cross validation across a sequence of INLA mesh `cutoff` values i.e., smaller cutoff = higher resolution mesh:

```{r tmb-sim-fits, cache=TRUE, message=FALSE, warning=FALSE}
mcv <- lapply(cutoffs, function(k) {
  cat("Cutoff:", k, "\n")
  mesh <- make_mesh(s, c("x", "y"), cutoff = k)
  cat("Mesh n:", mesh$mesh$n, "\n")
  set.seed(123)
  sdmTMB_cv(
    observed ~ 1 + x1,
    data = d,
    silent = TRUE,
    parallel = TRUE,
    # priors = sdmTMBpriors(matern_s = pc_matern(range_gt = 0.05, sigma_lt = 1)),
    mesh_args = list(cutoff = k, xy_cols = c("x", "y")),
    family = gaussian(),
    fold_ids = "fold_ids"
  )
})
```

```{r}
lls <- purrr::map_dbl(mcv, "sum_loglik")
# `vert` = vertices
vert <- purrr::map_dbl(mcv, function(x) x$models[[1]]$spde$mesh$n)
```

```{r}
plot(cutoffs, lls, type = "o")
plot(vert, lls, type = "o")
```

What about the 'in-bag' data for each fold?
  
```{r}
in_bag <- purrr::map_dfr(mcv, function(x) {
  nll <- purrr::map_dbl(seq_along(x$models), 
    function(i) x$models[[i]]$model$objective)
  tibble(ll = -nll, cv_id = seq_along(nll))
}, .id = "cutoff_id")
left_join(in_bag, tibble(cutoff_id = as.character(seq_along(vert)), 
  vertices = vert)) %>% 
  ggplot(aes(vertices, ll, group = cv_id, colour = as.factor(cv_id))) + 
  geom_line()
```

Extract spatial random fields and predictions for first fold.

```{r tmb-sim-preds, cache=TRUE}
nd <- expand.grid(x = seq(-1, 1, length.out = 150), 
  y = seq(-1, 1, length.out = 150))
nd$x1 <- s$x1
p <- purrr::map_dfr(seq_along(mcv), function(i) {
  p <- predict(mcv[[i]]$models[[1]], newdata = nd)
  p$cutoff <- cutoffs[i]
  p$vertices <- mcv[[i]]$models[[1]]$spde$mesh$n
  p$cv_log_lik <- mcv[[i]]$sum_loglik
  p
})
p <- group_by(p, cutoff) %>%
  mutate(ll_relative = cv_log_lik - max(p$cv_log_lik))
p$meta <- paste0("Vert.: ", stringr::str_pad(p$vertices, 3L), ", ",
  "CV LL: ", round(p$ll_relative, 2))
p$meta2 <- paste0("Cut.: ", stringr::str_pad(p$cutoff, 2L, "right"), ", ",
  "CV LL: ", round(p$ll_relative, 2))
```

Note the log likelihood (LL) is relative here so that 0 is the highest.

Predictions from 1st fold:

```{r, fig.width=8, fig.height=8}
ggplot(p, aes(x, y, fill = omega_s)) +
  geom_raster() +
  scale_fill_gradient2() +
  facet_wrap(~meta) +
  coord_fixed(expand = FALSE) +
  ggtitle("Estimated spatial random field")
```

Truth:

```{r}
ggplot(s, aes(x, y, fill = omega_s))+ 
  geom_raster() +
  scale_fill_gradient2() +
  coord_fixed(expand = FALSE)
```

Predictions from 1st fold:

```{r, fig.width=8, fig.height=8}
ggplot(p, aes(x, y, fill = est)) +
  geom_raster() +
  scale_fill_viridis_c() +
  facet_wrap(~meta) +
  coord_fixed(expand = FALSE) +
  ggtitle("Predicted value")
```

Truth:

```{r}
ggplot(s, aes(x, y, fill = mu)) +
  geom_raster() +
  scale_fill_viridis_c() +
  coord_fixed(expand = FALSE)
```


